---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
First, we will be loading the file content into data DataFrame, and then we will extract only data that has no NA values:
```{r echo = TRUE}
data <- read.csv(file="../activity.csv",head=TRUE,sep=",")
clean_data <- data[complete.cases(data),]
```


## What is mean total number of steps taken per day?
I will sum the steps value and group it by date. Then, a histogram graph is printed:
```{r echo = TRUE}
sum_by_date <- aggregate(clean_data$steps, by=list(date=clean_data$date), FUN=sum)
hist(sum_by_date$x, col="blue")
```

this is the Mean by date:
```{r echo = TRUE}
mean_by_date <- aggregate(clean_data$steps, by=list(date=clean_data$date), FUN=mean)
mean_by_date
```

And this is the Median by Date:
```{r echo = TRUE}
median_by_date <- aggregate(clean_data$steps, by=list(date=clean_data$date), FUN=median)
median_by_date
```


## What is the average daily activity pattern?

We will create mean for the steps per interval and plot the graph:
```{r echo = TRUE}
mean_by_interval <- aggregate(clean_data$steps, by=list(interval=clean_data$interval), FUN=mean)
plot(mean_by_interval$interval,mean_by_interval$x,type = "l",xlab="5-min Interval",ylab="Averaged Steps all Days")
```

The time interval that has the maximum number of steps all over the time interval is calculated like this:
```{r echo = TRUE}
max_interval <- mean_by_interval[mean_by_interval$x == max(mean_by_interval$x),]
```
Hence the The interval is: **`r max_interval$interval`**

## Imputing missing values
1. Since I already got the clean data without NA, I just need to calculate the diffirence between the number of rows as follows:

```{r echo = TRUE}
nrow(data)- nrow(clean_data)
```

2. To fill the data set, I will fill the NA data with the mean for that time interval.

```{r echo = TRUE}
na_data = data[is.na(data),]
for(i in 1:nrow(na_data))
{
        na_data[i,"steps"] = mean_by_interval[mean_by_interval$interval == na_data[i,"interval"],]$x
}
```

After filling the na_data, we will be having 2 data frames na data frame and non-na data frame (which we called it clean_data) 

3. To create the new data set, we will be appending na_data to clean_data using the below code:
```{r echo = TRUE}
new_data <- rbind(na_data, clean_data)
head(new_data)
```

4. Lets check out the new data how it will look like and how it will differ in terms of mean and median:
```{r echo = TRUE}
sum_by_date_new <- aggregate(new_data$steps, by=list(date=new_data$date), FUN=sum)
hist(sum_by_date_new$x, col="blue")
```

this is the Mean by date:
```{r echo = TRUE}
mean_by_date_new <- aggregate(new_data$steps, by=list(date=new_data$date), FUN=mean)
mean_by_date_new
```

And this is the Median by Date:
```{r echo = TRUE}
median_by_date_new <- aggregate(new_data$steps, by=list(date=new_data$date), FUN=median)
median_by_date_new
```



## Are there differences in activity patterns between weekdays and weekends?

1. We will create a new valiable called day_type which has weekend or weekday values:
```{r echo = TRUE}
for(i in 1:nrow(new_data))
{
        if(!(weekdays(as.Date(new_data[i,]$date)) %in% c('Saturday','Sunday')))
        {
                type <- "weekday"
        }
        else
        {
                type <- "weekend"
        }
        new_data[i,"day_type"] = type
}

```

2. We need to group the new_data by day_type and interval and then plot the result:
```{r echo = TRUE}
mean_by_day_type_interval <- aggregate(new_data$steps, by=list(interval=new_data$interval,day_type=new_data$day_type), FUN=mean)

library(ggplot2)
g <- ggplot(mean_by_day_type_interval, aes(interval,  x))
g + geom_line() +  facet_grid(day_type ~ .) + ggtitle("Avg. Steps per day interval for Weekend and Weekday")

```